---
title: "Coral MuSE"
author: "LMM, EM, MV, AC, et al"
date: "14 de enero de 2019"
output: 
  html_document:
    code_download: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages

```{r packages}
library(tidyverse)
library(magrittr)
library(vegan)
#library(cowplot)
library(jtools)
library(trafo)
library(colorblindr)
library(cowplot)
library(ggrepel)
source('r/MSEgroup.r')
source('r/BootResampler.R')
```

```{r}
if (!dir.exists('analysis/figures')){dir.create('analysis/figures')}

if (!dir.exists('analysis/outcomes')){dir.create('analysis/outcomes')}
```


# Field data

```{r field dataset}
reef_pts <- read.csv('data/Points_Transect_BenthicSubstrate.csv', 
                     check.names = F, stringsAsFactors = T)
#
groups <- reef_pts[,2:6]
#
available_files <- reef_pts[reef_pts$File %in%
                        list.files("data/transects"), ]

coral_spp <- read.csv('data/caribbean_spp.csv')
```

```{r coral absolute cover, warning=FALSE}
coral_cover <-   reef_pts %>% 
  select(-c(1:6)) %>% 
  select(one_of(as.character(coral_spp[coral_spp$Group.ID %in% c(1:9), "Species.name"]))) %>% 
  divide_by(375) %>% 
  multiply_by(100) %>%
  mutate(dummy = 1) %>%   #add a dummy variable
  vegdist(method = "bray")
```

```{r mse and tidy format}
raw_mse <- coral_cover %>% 
  MSEgroup.d(groups$Site) %>% 
  data.frame() %>% 
  mutate(sites = rownames(.)) %>% 
  gather(-sites, key = 'replicate', value = "MSE") %>% 
  separate(replicate, 
           into = c('value', 'replicate'), 
           sep = '[.]') %>% 
  spread(value, MSE) %>% 
  filter(replicate != 1) %>% #exclude 1st transect
  mutate(replicate = as.numeric(replicate), 
         sites =  as.factor(sites)) %>% 
  rename(mse = means) %>% 
  filter(sites != 'Petaquire' | replicate != 4) %>% 
  mutate(replicate = replace(replicate, 
                             sites == 'Petaquire' & replicate == 3, 4)
         )
```

```{r}
field_mse <- raw_mse %>% 
  filter(replicate == 4) %>% 
  left_join(groups[groups$Transect == 't3',], 
            by = c('sites' = 'Site')) %>% 
  mutate(range = upper-lower) %>% 
  mutate(prop = range/mse) %>% 
  arrange(Locality)

field_mse$sites <- factor(field_mse$sites, 
                           levels = field_mse$sites[order(field_mse$Locality)])
```


# Achieved precision

```{r fig01}
ggplot(field_mse, aes(x = fct_rev(sites), 
                      y = mse, 
                      fill = Locality)) + 
  geom_bar(stat = 'identity') +
  geom_errorbar(aes(ymin = lower, ymax = upper), 
                width= .1) +
  coord_flip() +
  geom_hline(yintercept = mean(field_mse$mse),
             alpha=0.25,
             size = 2) +
  labs(x = "Sites", y = 'Multivariate standard error') +
  scale_fill_OkabeIto()

  #ggsave('analysis/figures/fig01.svg')
```


```{r coral richness, warning=FALSE}
coral_rich <- reef_pts %>% 
  select(-c(1:6)) %>% 
  select(one_of(as.character(coral_spp[coral_spp$Group.ID %in% c(1:9), "Species.name"]))) %>%
  specnumber(groups = reef_pts$Site) %>% 
  data.frame() %>% 
  rename(richness = '.') %>% 
  rownames_to_column('sites')
```

# Univariate precision

```{r}
cover_prec <- reef_pts %>% 
  #select(-c(1:6)) %>% 
  select(one_of(as.character(coral_spp[coral_spp$Group.ID %in% c(1:9), "Species.name"]))) %>% 
  divide_by(375) %>% 
  multiply_by(100) %>%
  mutate(total = rowSums(.[])) %>%
  select(total) %>% 
  bind_cols(reef_pts[,c(1:6)]) %>% 
  group_by(Site, Locality) %>% 
  summarise(mean_cover = mean(total), sd_cover = sd(total), n = length(total)) %>% 
  mutate(precision = (sd_cover/sqrt(n))/mean_cover) %>% 
  mutate(se = (sd_cover/sqrt(n))) %>% 
  rename(sites = Site)

cover_prec$sites <- factor(cover_prec$sites, 
                           levels = cover_prec$sites[order(cover_prec$Locality)])
```

```{r coral summary}
coral_summ <- field_mse[,c('mse', 'sites', 'Locality')] %>% 
  left_join(cover_prec[,c('se', 'sites', 'Locality')]) %>% 
  left_join(coral_rich)
```


```{r fig02}
p1 <- ggplot(coral_summ, aes(x = se,
                             y = mse, 
                             label = sites,
                             color = Locality,
                             size = 2)) +
  geom_point(position = 'jitter') +
  scale_color_OkabeIto() +
  guides(size = FALSE) +
  labs(x = "Univariate standard error",
       y = "Multivariate standard error")+
  annotate("text", 
           label = paste("Pearson's correlation =", 
                                 round(cor(coral_summ$mse, coral_summ$se), 2)), 
           x = 3, y = 0.25)

p2 <- ggplot(coral_summ, aes(x = richness, 
             y = mse, 
             label = sites,
             color = Locality,
             size = 2)) +
  geom_point(position = 'jitter') +
  scale_color_OkabeIto() +
  guides(size = FALSE) +
  labs(x = "Coral species richness",
       y = "Multivariate standard error") +
  annotate("text", 
           label = paste("Pearson's correlation =", 
                                 round(cor(coral_summ$mse, coral_summ$richness), 2)), 
           x = 8.5, y = 0.25)

plot_grid(
  plot_grid(
    p1 + theme(legend.position = 'none'),
    p2 + theme(legend.position = 'none'),
    nrow = 1, align = "v", labels = c('A', 'B')
    ),
  get_legend(p1 + scale_shape(guide = FALSE) + 
      theme(legend.position = "bottom")),
  ncol=1, rel_heights=c(.85, .15)
)

#ggsave('analysis/figures/fig02.svg')
```

Now we choose the set of sites for resampling

```{r}
field_mse %>% 
  arrange(desc(mse)) %>% 
  filter(range %in% head(range, 12))
```


# Transects


#Dataset construction

```{r}
NTransects <- 20
NQuadrats <- seq(5, 25, 5)
NPoints <- seq(25, 75, 25)

DF <- NULL

for(i in 1:length(NTransects)) {
  for(j in 1:length(NQuadrats)) {
    for(k in 1:length(NPoints)) {
      
      boot_DF <- BootResampler(DataFrame = available_files, 
                    nbTransects = NTransects[i], 
                    nbQuadrats = NQuadrats[j], 
                    nbPoints = NPoints[k], 
                    transects_folder = "data/transects/", 
                    SpMatrix_Path = coral_spp, 
                    randomseed = 1)
      
      group_DF <- boot_DF %>% 
        select(2:6)
      
      mse_DF <- boot_DF %>% 
        select(-c(1:6)) %>%
        select(one_of(as.character(coral_spp[coral_spp$Group.ID %in% c(1:9), "Species.name"]))) %>% 
        divide_by(375) %>% 
        multiply_by(100) %>%
        mutate(dummy = 1) %>%   #add a dummy variable
        vegdist(method = "bray") %>% 
        MSEgroup.d(group_DF$Site) %>% 
        data.frame() %>% 
        mutate(sites = rownames(.)) %>% 
        gather(-sites, key = 'replicate', 
               value = "MSE") %>% 
        separate(replicate, 
                 into = c('value', 'replicate'), 
                 sep = '[.]') %>% 
        spread(value, MSE) %>% 
        filter(replicate != 1) %>% #exclude 1st transect
        mutate(replicate = as.numeric(replicate), 
               sites =  as.factor(sites)) %>% 
        rename(mse = means) %>% 
        mutate(ntransects = NTransects[i], 
               nquads = NQuadrats[j], 
               npoints = NPoints[k])
      
      DF <- bind_rows(DF, mse_DF)
      
      print(paste("Transects: ", NTransects[i], 
                  "| Quadrats: ", NQuadrats[j], 
                  "| Points: ", NPoints[k]))
      
    }
  }
}
```

```{r}
DF <- DF %>%
  mutate(range = upper-lower) %>% 
  mutate(prop = range/mse) %>% 
  rename(transects = replicate, quadrats = nquads, points = npoints)
```

We can write this object to save computation time in the future:

```{r}
#saveRDS(DF, 'analysis/outcomes/DF.rds')
#DF <- readRDS('analysis/outcomes/DF.rds')
```


#Field vs bootstrapped data

We can compare if the MuSE behaves the same in the field and bootstrapped datasets

```{r comparison}
set1 <- raw_mse %>% 
          select(mse, sites, replicate) %>%
          rename(transects = replicate) %>% 
          mutate(type='field')

set2 <- DF %>% 
  select(mse, sites, replicate) %>% 
  rename(transects = replicate) %>% 
  mutate(type='boot')
  
bind_rows(set1, set2) %>% 
  filter(sites =='Salinas') %>% #choose an specific site
  ggplot(aes(x = transects, y = mse, shape = type, color = type)) + 
  geom_point(position = 'jitter')
```

# Plots

##  Global

```{r sites as replicates}
ggplot(DF, aes(x = transects, y = mse)) +
  geom_point() +
  facet_grid(quadrats~points) +
  theme(panel.grid.major = element_line(colour='gray90'))
```

##  Particular cases

```{r Case Punta Cruz}
DF %>% 
  filter(sites %in% c('Salinas'))%>% 
  mutate(points = as.factor(points), 
         quadrats = as.factor(quadrats)) %>% 
  ggplot(aes(x = transects, y = mse)) +
  geom_point() +
  geom_ribbon(aes(ymin = lower, ymax = upper), 
              linetype = 2, 
              alpha = 0.1) +
  facet_grid(quadrats~points) +
  theme(panel.grid.major = element_line(colour='gray90'))
```

# Regression

```{r regression}
mod <- lm(mse~points*quadrats*transects, data = DF)

#plot(mod)
```

```{r}
diagnostics(trafo_lm(mod))

bc_mod <- trafo_lm(mod)

summary(bc_mod$trafo_mod)
```

```{r figure03}
plot_coefs(bc_mod$trafo_mod, colors = 'black', scale = TRUE) +
  theme_nice() +
  geom_rect(xmin = -Inf, xmax = Inf, ymin = 0.6, ymax = 1.4,   
            fill = "gray", alpha = 0.05) +
  geom_rect(xmin = -Inf, xmax = Inf, ymin = 1.6, ymax = 4.4,   
            fill = "gray", alpha = 0.05) +
  geom_rect(xmin = -Inf, xmax = Inf, ymin = 4.6, ymax = 7.4,  
            fill = "gray", alpha = 0.05) +
  labs(y = 'Term')
#ggsave('analysis/figures/fig03.svg')
```

# New configuration

```{r}
boot_DF2 <- BootResampler(DataFrame = available_files, 
                    nbTransects = 10, 
                    nbQuadrats = 10, 
                    nbPoints = 25, 
                    transects_folder = "data/transects/", 
                    SpMatrix_Path = coral_spp, 
                    randomseed = 1)

group_DF2 <- boot_DF2 %>% 
        select(2:6)
      
boot_disp <- boot_DF2 %>% 
  select(-c(1:6)) %>%
  select(one_of(as.character(coral_spp[coral_spp$Group.ID %in% c(1:9), "Species.name"]))) %>% 
  divide_by(375) %>% 
  multiply_by(100) %>%
  mutate(dummy = 1) %>%   #add a dummy variable
  vegdist(method = "bray") %>% 
  betadisper(group = group_DF2$Site, type = 'centroid')

boot_vectors <- boot_disp$vectors %>% 
  data.frame() %>% 
  mutate(Site = as.factor(group_DF2$Site))

boot_centroids <- boot_disp$centroids %>% 
  data.frame() %>% 
  rownames_to_column('Site')
```

## Field subset

```{r}
sites_subset <- reef_pts %>% 
  filter(Site %in% droplevels(group_DF2$Site)) %>% 
  select(2:6)

coral_subset <- reef_pts %>% 
  filter(Site %in% droplevels(group_DF2$Site)) %>%
  select(-c(1:6)) %>% 
  select(one_of(as.character(coral_spp[coral_spp$Group.ID %in% c(1:9), "Species.name"]))) %>% 
  divide_by(375) %>% 
  multiply_by(100) %>%
  mutate(dummy = 1) %>%   #add a dummy variable
  vegdist(method = "bray") %>% 
  betadisper(group = sites_subset$Site, type = 'centroid')

coral_vectors <- coral_subset$vectors %>% 
  data.frame() %>% 
  mutate(Site = droplevels(sites_subset$Site)) %>% 
  select(c(PCoA1, PCoA2, Site))

coral_centroids <- coral_subset$centroids %>% 
  data.frame() %>% 
  rownames_to_column('Site')
```

## both pco

```{r}
p1 <- ggplot(coral_centroids, aes(x = PCoA1, y = PCoA2, 
                                  label = Site, color = Site)) +
  geom_point() +
  geom_label_repel(aes(label = Site), point.padding = 2) +
  ylim(-0.7, 0.7) +
  xlim(-0.7, 0.7) +
  ggtitle('15q, 4t') +
  theme(legend.position = 'none') +
  stat_ellipse(data = subset(coral_vectors, Site %in% c('Salinas')), 
               alpha = 0.3,
               level = 0.68) +
  stat_ellipse(data = subset(coral_vectors, Site %in% c('Charagato')), 
               alpha = 0.3,
               level = 0.68) +
  stat_ellipse(data = subset(coral_vectors, Site %in% c('Punta Conejo')), 
               alpha = 0.3,
               level = 0.68) +
  stat_ellipse(data = subset(coral_vectors, Site %in% c('Cienaga oeste')), 
               alpha = 0.3,
               level = 0.68) +
  stat_ellipse(data = subset(coral_vectors, Site %in% c('Punta Cruz')), 
               alpha = 0.3,
               level = 0.68) +
  stat_ellipse(data = subset(coral_vectors, Site %in% c('Guabinitas')), 
               alpha = 0.3,
               level = 0.68) +
  stat_ellipse(data = subset(coral_vectors, Site %in% c('Madrisqui')), 
               alpha = 0.3,
               level = 0.68) +
  stat_ellipse(data = subset(coral_vectors, Site %in% c('Mero')), 
               alpha = 0.3,
               level = 0.68) +
    stat_ellipse(data = subset(coral_vectors, Site %in% c('Punta de Media Legua')), 
               alpha = 0.3,
               level = 0.68) +
    stat_ellipse(data = subset(coral_vectors, Site %in% c('Salinas')), 
               alpha = 0.3,
               level = 0.68) +
  stat_ellipse(data = subset(coral_vectors, Site %in% c('Dos Mosquises Sur')), 
               alpha = 0.3,
               level = 0.68) +
  stat_ellipse(data = subset(coral_vectors, Site %in% c('Boca Seca')), 
               alpha = 0.3,
               level = 0.68) +
  stat_ellipse(data = subset(coral_vectors, Site %in% c('La Muerta')), 
               alpha = 0.3,
               level = 0.68)
```

```{r}
p2 <- ggplot(boot_centroids, aes(x = PCoA1, y = -PCoA2, label = Site)) +
  geom_point() +
  geom_label_repel(aes(label = Site), point.padding = 0.5) +
  ylim(-0.7, 0.7) +
  xlim(-0.7, 0.7) +
  ggtitle('10q, 10t') +
  stat_ellipse(data = subset(boot_vectors, Site %in% c('Salinas')), 
               alpha = 0.3,
               level = 0.68) +
    stat_ellipse(data = subset(boot_vectors, Site %in% c('Charagato')), 
               alpha = 0.3,
               level = 0.68) +
    stat_ellipse(data = subset(boot_vectors, Site %in% c('Punta Conejo')), 
               alpha = 0.3,
               level = 0.68) +
    stat_ellipse(data = subset(boot_vectors, Site %in% c('Cienaga oeste')), 
               alpha = 0.3,
               level = 0.68)
```

```{r}
plot_grid(p1,p2)
```

```{r}
ggplot(aes(x = PCoA1, y = PCoA2, label = Site), data = subset(coral_centroids, Site %in% c('Cienaga oeste'))) +
  geom_point(shape = 2) +
  geom_text(aes(x = PCoA1, y = PCoA2), data = subset(coral_vectors, Site %in% c('Cienaga oeste'))) +
  stat_ellipse(data = subset(coral_vectors, Site %in% c('Cienaga oeste')), alpha = 0.3,
               level = 0.68) +
  geom_point(aes(x = PCoA1, y = PCoA2), data = subset(boot_vectors, Site %in% c('Cienaga oeste'))) +
  geom_point(aes(x = PCoA1, y = PCoA2), shape = 2,
             data = subset(boot_centroids, Site %in% c('Cienaga oeste'))) +
  stat_ellipse(data = subset(boot_vectors, Site %in% c('Cienaga oeste')), alpha = 0.3,
               level = 0.68)
```



